---
title: "Trends in Nighttime Lights Across Admin Zones"
author: ''
date: ''
format:
  html:
    toc: true
    code-fold: false
knitr:
  opts_chunk:
    warning: false
    message: false
---

This page shows how to examine trends in nighttime lights across administrative zones for a country. The page illustrates using nighttime lights data from [NASA Black Marble](https://blackmarble.gsfc.nasa.gov/) using the [BlackMarbleR](https://github.com/worldbank/blackmarbler) R package, where the package facilitates querying data from the [NASA LAADS Archive](https://ladsweb.modaps.eosdis.nasa.gov/).

# Setup

Querying data from the NASA LAADS archive---and using the BlackMarbleR package---requires a NASA bearer token. To obtain a token, follow [these steps](https://github.com/worldbank/blackmarbler?tab=readme-ov-file#bearer-token-).

# Code

* ADD LINK TO COLAB

```{r}
#| include: false
library(tidyverse)
bearer <- read_csv("~/Dropbox/bearer_bm.csv") %>%
  pull(token)
```


```{r}
## Load packages
library(tidyverse)    # R workhorse
library(sf)           # Working with spatial data
library(geodata)      # To query country boundary data from GADM
library(blackmarbler) # To query Black Marble NTL data
library(kableExtra)   # Making tables

## Bearer Token
bearer <- "BEARER TOKEN HERE"
```

```{r}
#| include: false
bearer <- read_csv("~/Dropbox/bearer_bm.csv") %>%
  pull(token)
```

```{r}
## Get polygons of ADM1 for example country (Rwanda)
roi_sf <- gadm('RWA', level=1, path = tempdir()) %>%
  st_as_sf()

## Extract total annual nighttime lights within Rwanda's ADM1 units
ntl_df <- bm_extract(roi_sf = roi_sf,
                     product_id = "VNP46A4",
                     date = 2012:2023,
                     bearer = bearer,
                     aggregation_fun = "sum",
                     quiet = T)

## Examine data
ntl_df %>%
  select(COUNTRY, NAME_1, date, contains("n_")) %>%
  head() %>%
  kable()

## Plot trends
ntl_df %>%
  ggplot() +
  geom_col(aes(x = date, 
               y = ntl_sum),
           fill = "darkorange") +
  facet_wrap(~NAME_1) +
  labs(x = NULL,
       y = "NTL",
       title = "Annual Nighttime Lights Across Rwanda's ADM1s") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```

# Related Examples

* NTL trends using BlackMarblePy
* NTL trends using GEE and Python
* NTL trends using GEE and R
* NTL trends using GEE code editor

# References

* World Bank economic monitors, including monitors for [Lebanon](https://datapartnership.org/lebanon-economic-monitor/notebooks/ntl-analysis/README.html), [Myanmar](https://datapartnership.org/myanmar-economic-monitor/notebooks/nighttime-lights/analysis-2023/README.html)




